---
title: "STAT652 Project 5. Neural network"
author: "Chun Yin Kong"
date: "3/8/2021"
output: pdf_document
---

# Initial Setup
```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(tidymodels)
library(mdsr)
library(parallel)
library(remotes) # For installing traversc/trqwe
library(trqwe) #parallel 1
library(tictoc)
library(future) # parallel processing 2
```

# Step 1 Collect Data
## Reading Data
```{r warning=FALSE, message=FALSE, eval=FALSE}
tic()
accepted_data <- mcreadRDS(file="data/accepted_data.rds", mc.cores = min(parallel::detectCores(), 16))
# Requires "pigz" on ubuntu
toc()
```

## Analyzing data
```{r}
glimpse(accepted_data)
```

```{r}
nrow(accepted_data)
ncol(accepted_data)
```

# Step 2 Exploring and Preparing the data
Creating 75-25 Split of Training and Testing Data
```{r}
set.seed(123123)
n <- nrow(accepted_data)
accepted_data_part <- accepted_data %>%
  select(-id, -member_id) %>% #removing id variables
  mutate(loan_status = as.factor(loan_status)) %>%
  filter(str_detect(issue_d, "2012|2013|2014")) %>%
  select(loan_status, where(is.numeric)) %>% # letting all numeric variables remain, others all drop
  initial_split(prop=0.75)

train <- accepted_data_part %>%
  training()

test <- accepted_data_part %>%
  testing()

list(train, test) %>%
  map_int(nrow)

remove(accepted_data, accepted_data_part) # removing large dataset to avoid RAM shortage
```

# Step 3 Training a model on the data
Setting formula for Neural Network
```{r}
form <- as.formula("loan_status ~ loan_amnt + funded_amnt + funded_amnt_inv + int_rate + installment + annual_inc + delinq_2yrs + fico_range_low + fico_range_high")
```

```{r warning=FALSE, message=FALSE}
tic()
mod_nn <-mlp(epochs = 100, mode="classification", hidden = 5) %>%
  set_engine("keras") %>%
  fit(form, data=train)
toc()
```

# Step 4 Evaluating Model Performance
```{r}
pred <- pred %>%
  bind_cols(
    predict(mod_nn, new_data = test, type = "class")
  ) %>%
  rename(loan_status_nn = .pred_class)
```

```{r}
confusion_nn <- pred %>%
  conf_mat(truth = loan_status, estimate = loan_status_nn)

confusion_nn
```

```{r}
accuracy(pred, loan_status, loan_status_nn)
```

# Step 5 Improving Model Performance
We can improve the neural network model by adjusting the hidden layers of the neural network, as well as the formula to be fit in the neural network.