---
title: "STAT652 Project 7. kNN"
author: "Chun Yin Kong"
date: "3/8/2021"
output: html_document
---

# Initial Setup
```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(tidymodels)
library(mdsr)
library(parallel)
library(remotes) # For installing traversc/trqwe
library(trqwe) #parallel 1
library(tictoc)
library(future) # parallel processing 2
```

# Step 1 Collect Data
## Reading Data
```{r warning=FALSE, message=FALSE, eval=FALSE}
tic()
accepted_data <- mcreadRDS(file="data/accepted_data.rds", mc.cores = min(parallel::detectCores(), 16))
# Requires "pigz" on ubuntu
toc()
```

## Analyzing data
```{r}
glimpse(accepted_data)
```

```{r}
nrow(accepted_data)
ncol(accepted_data)
```

# Step 2 Exploring and Preparing the data
Creating 75-25 Split of Training and Testing Data
```{r warning=FALSE, message=FALSE}
set.seed(123123)
n <- nrow(accepted_data)
accepted_data_part <- accepted_data %>%
  select(-id, -member_id) %>% #removing id variables
  mutate(loan_status = as.factor(loan_status)) %>%
  filter(str_detect(issue_d, "2012|2013|2014")) %>%
  select(loan_status, where(is.numeric)) %>% # letting all numeric variables remain, others all drop
  initial_split(prop=0.75)

train <- accepted_data_part %>%
  training()

test <- accepted_data_part %>%
  testing()

list(train, test) %>%
  map_int(nrow)

remove(accepted_data, accepted_data_part) # removing large dataset to avoid RAM shortage
```
# Step 3 Training a model on the data

```{r}
form <- as.formula("loan_status ~ loan_amnt + funded_amnt + funded_amnt_inv + int_rate + installment + annual_inc + delinq_2yrs + fico_range_low + fico_range_high")
```

```{r}
tic()
train_q <- train %>%
  select(loan_status, where(is.numeric)) 

mod_knn <- nearest_neighbor(neighbors = 5, mode = "classification") %>%
  set_engine("kknn", scale=TRUE) %>%
  fit(form, data = train_q)
toc()
```

# Step 4 Evaluating Model Performance
```{r}
pred <- pred %>%
  bind_cols(
    predict(mod_knn, new_data = test, type = "class")
  ) %>%
    rename(loan_status_knn = .pred_class)
```

```{r}
confusion_knn <- pred %>%
  conf_mat(truth = loan_status, estimate = loan_status_knn)

confusion_knn
```

```{r}
accuracy(pred, loan_status, loan_status_knn)
```

# Step 5 Improving Model Performance
We can improve the kNN model by changing the number of k, the number of nearest neighbor that used to classify to improve the performance. Sometimes a small number of k could lead to overfitting and at the same time a higher number of k could lead to underfitting.